---
title: "Plot Grid Search results"
author: "Sihan Chen, Richard Futrell, Kyle Mahowald"
date: '2022-09-28'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(ggrepel)
library(RColorBrewer)
library(ggnewscale)
library(latex2exp)
library(gridExtra)
library(brms)
library(ggpubr)
library(ggrepel)
library(tidybayes)
library(patchwork)
```

This script generates figures from Experiment 2, where we try different free parameter setting and see how the efficiency of real lexicon changes with respect to that of simulated lexicons.


Read dataset: 
```{r}
d = read.csv('sheets/total_search_results.csv') %>%
  mutate(favoring = ifelse(abs(source) > abs(goal), 'goal', ifelse(abs(source) < abs(goal), 'source', 'neither')),
         place_location = ifelse(goal < 0 & source > 0 | goal > 0 & source < 0, 'PLACE central', 'PLACE marginal')) %>%
  mutate(favoring = factor(favoring, levels = c('goal', 'source', 'neither')),
         place_location = factor(place_location)) %>% 
  rowwise() %>% mutate(cat = paste(favoring, place_location, sep = '_')) %>% ungroup() %>%
  mutate(code = case_when(cat == 'goal_PLACE central' ~ 'S--P-G',
                           cat == 'goal_PLACE marginal' ~  'S-G-P',
                           cat == 'neither_PLACE central' ~ 'S-P-G',
                           cat == 'neither_PLACE marginal' ~ 'S=G-P',
                           cat == 'source_PLACE central' ~ 'S-P--G',
                           cat == 'source_PLACE marginal' ~ 'P-S-G'),
         code = factor(code, levels = c('S--P-G', 'S-G-P', 'S-P-G', 'S=G-P', 'S-P--G','P-S-G'))) %>%
  group_by(place_location) %>% ungroup()

```

This is a scary plot: we might put it in the appendix

```{r}
ggplot(d, aes(x = goal, y = source, size = mean_gNID)) +
  facet_wrap(~prior_spec) +
  geom_point() +
  geom_point(data = d %>% slice_min(order_by = mean_gNID, n = 1), color = 'red') +
  geom_abline(slope = -1, intercept = 0, color = 'green') +
  theme_bw(20)

```


Plot mean gNID under each prior configuration
```{r}
prior_specs <- unique(d$prior_spec[d$prior_spec != 'place_goal_source'])
my_comparisons <- list(c("place_goal_source", prior_specs[1]), 
                       c("place_goal_source", prior_specs[2]), 
                       c("place_goal_source", prior_specs[3]), 
                       c("place_goal_source", prior_specs[4]), 
                       c("place_goal_source", prior_specs[5]), 
                       c("place_goal_source", prior_specs[6]), 
                       c("place_goal_source", prior_specs[7]))
ggplot(d, aes(x = reorder(prior_spec, mean_gNID), y = mean_gNID)) +
#facet_wrap(~code) +
stat_summary(geom = 'col', fun = mean) +
scale_fill_manual(values = c('gray90', 'gray90', 'gray90','gray90','gray90', 'gray90', 'gray50', 'gray90')) +
stat_summary(geom = 'errorbar', fun.data = mean_cl_boot) +
geom_jitter(width = 0.05, height = 0, alpha = 0.5) +
theme_classic(14) +
theme(axis.text.x = element_text(angle = 30, vjust=0.5)) +
stat_compare_means(method='t.test', comparisons=my_comparisons) +
xlab('Need distribution (from most frequent to least frequent)') +
ylab('Average gNID')

ggsave('figures/paper_figures/need_distribution.png')

d %>% group_by(prior_spec) %>% summarise(mean = mean(mean_gNID)) %>% ungroup()
```


Plot mean gNID under each pgs representation cost
```{r}
codes <- as.character(unique(d$code[d$code != 'S--P-G']))
my_comparisons <- list(c('S--P-G', codes[1]),
                       c('S--P-G', codes[2]),
                       c('S--P-G', codes[3]),
                       c('S--P-G', codes[4]),
                       c('S--P-G', codes[5]))

ggplot(d, aes(x = reorder(code, mean_gNID), y = mean_gNID, fill = code)) +
#facet_wrap(~prior_spec) +
stat_summary(geom = 'col', fun = 'mean', width = 0.3) +
geom_jitter(width = 0.05, height = 0, alpha = 0.5) +
stat_summary(geom = 'errorbar', fun.data = 'mean_cl_boot', width = 0.3) +
theme_classic(25) +
stat_compare_means(method='t.test', comparisons=my_comparisons) +
xlab('Place-Goal-Source coordinates') +
ylab('Average gNID') +
theme(panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.title.x=element_blank(),
      #axis.text.x=element_blank(),
      legend.position = "none"
      ) +
scale_fill_manual(values = c('gray50', 'gray90', 'gray90', 'gray90','gray90','gray90'))

ggsave('figures/paper_figures/rep_cost.png')

```

Stat: which factor matters more
```{r}
m <- brm(mean_gNID ~ (1|prior_spec) + (1|code), data =d)
# m <- brm(mean_gNID ~ prior_spec + code, data = d,
#          family=gaussian(),
#          warmup=2000,
#          iter=10000,
#          chain=4,
#          thin=10)
# 
# m %>% summary()

#hypothesis(m, 'abs(prior_specplace_source_goal) < abs(codeSMGMP)')

# do BRMS
# use the best cost function and the worst cost function, best prior and the worst prior
p1 <- m %>% spread_draws(r_prior_spec[prior, ], r_code[code, ]) %>% 
  ggplot(aes(x=r_code, y=code)) + 
  geom_halfeyeh() + 
  xlim(-0.3, 0.3) + 
  theme_bw(15) +
  ylab('orientation confusion cost') +
  xlab('intercept')
p2 <- m %>% spread_draws(r_prior_spec[prior, ], r_code[code, ]) %>% 
  ggplot(aes(x=r_prior_spec, y=prior)) + 
  geom_halfeyeh()+ 
  xlim(-0.3, 0.3) + 
  theme_bw(15) +
  ylab('need distribution permutation') +
  xlab('intercept')
(p1/p2) 



ggplot(m %>% spread_draws(r_prior_spec[prior, ], r_code[code, ]) %>% filter(abs(r_prior_spec) < abs(r_code)),
       aes(x=r_code, y=r_prior_spec)) + 
facet_grid(rows = vars(prior), cols = vars(code)) +
geom_point(color='lightgreen') + 
geom_point(data=m %>% spread_draws(r_prior_spec[prior, ], r_code[code, ]) %>% filter(abs(r_prior_spec) >= abs(r_code)),
           aes(x=r_code, y=r_prior_spec)) +
geom_abline(slope = 1, linetype='dashed', color='darkblue') +
geom_abline(slope= -1, linetype='dashed', color='darkblue') +
xlim(-0.3, 0.3) + 
ylim(-0.3, 0.3) +
coord_fixed() + 
theme_bw(15) +
theme(strip.text.y = element_text(angle=0)) +
xlab('random intercept of orientation confusion cost') +
ylab('random intercept of need distribution permutation')

ggsave('figures/paper_figures/comb.png')
m %>% spread_draws(r_prior_spec[prior, ], r_code[code, ]) %>% filter(abs(r_prior_spec) < abs(r_code)) %>% group_by(code, prior) %>%
  summarise(n1=n()) %>% left_join(
m %>% spread_draws(r_prior_spec[prior, ], r_code[code, ]) %>% filter(abs(r_prior_spec) >= abs(r_code)) %>% group_by(code, prior) %>%
  summarise(n2=n()), by = c('code', 'prior')) %>% 
  mutate(ratio=n1/n2) %>% View()
```



Plot gNID and mu
```{r}
d_mu <- read.csv('sheets/mu_search_gridsearch.csv')

ggplot(d_mu, aes(x = mu, y = mean_gNID)) +
  geom_line() +
  geom_point() +
  theme_classic(25) +
  xlab('decay parameter') +
  ylab('Average gNID')

```


Compare gNID and epsilon
```{r}
ggplot(d, aes(mean_gNID, mean_epsilon)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  theme_classic(25) +
  xlab('Mean gNID') +
  ylab(TeX('Mean $\\epsilon$'))
```

