---
title: "Map"
author: "Sihan Chen, Richard Futrell, Kyle Mahowald"
date: '2022-12-28'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(cowplot)
library(googleway)
library(sf)
library(ggrepel)
library(tidyverse)
library(rnaturalearth)
library(paletteer)
library(rnaturalearthdata)
library(tmap)
```

This script takes in the file called language_info.csv and makes a map of the real languages whose spatial deictic systems are studied.
The languages are taken from Nintemann et al. (2020). About half of the languages have a direct match in Glottolog with geographical coordinates.
For those languages that do not have a direct coordinate match in Glottolog (usually either a dialect or a collective term), we find the closest variety that has a coordinate.

Read the dataframe and set up a world map
```{r}
d <- read.csv('language_info.csv')
world <- ne_countries(scale = 'medium', returnclass = 'sf')
```


Fill in the points (Figure 4)
```{r}
sites <- st_as_sf(d, coords = c("longitude", "latitude"), crs = 4326, agr = 'constant') 
clrs <- paletteer_d("palettesForR::LaTeX", 68)
ggplot(data = world) +
  theme_bw() +
  geom_sf() +
  geom_sf(data = sites, aes(fill=family_name), size = 2, shape=22) +
  scale_fill_manual(values=clrs) +
  theme(legend.position = 'bottom', 
        legend.title = element_blank(),
        text = element_text(size=25))

ggsave('figures/orig_figures/Figure 4.png', width=14, height=10, unit='in')

```

